---
title: "Candidates Sentiment analysis"
author: "Daniel Robles"
date: "6/2/2020"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(include = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      comment = NA, # Remove "##" from results
                      fig.width=16, fig.height=9)
```

Check github [here](https://github.com/jboscomendoza/rpubs/tree/master/sentimientos_afinn)

```{r}
library(vctrs)
library(tidyverse)
library(tidytext)
library(tm)
library(lubridate)
library(zoo)
library(scales)
```

Load candidates tweets to develop sentiment analysis

```{r}
tuits <- read.csv("tuits_candidatos.csv", stringsAsFactors = F, fileEncoding = "latin1") %>%
  tbl_df()
head(tuits,3)
```

Load the lexicon to make the sentiment analysis

```{r}
afinn <- read.csv("lexico_afinn.en.es.csv", stringsAsFactors = F, fileEncoding = "latin1") %>%
  tbl_df()
head(afinn,3)
```

Separate date/hour in column "created_at", then "date. Take 2018 tweets only.

```{r}
tuits <-
  tuits %>%
  separate(created_at, into = c("Fecha", "Hora"), sep = " ") %>%
  separate(Fecha, into = c("Dia", "Mes", "Periodo"), sep = "/",
           remove = FALSE) %>%
  mutate(Fecha = dmy(Fecha), # Create column for date with lubridate
         Semana = week(Fecha) %>% as.factor(), # Create column for week of the year
         text = tolower(text)) %>% # Convert text column to lowercase
  filter(Periodo == 2018) # Take 2018 year only
head(tuits,3)
```

Inner join among `tuits` and `afinn` lexicon

```{r}
tuits_afinn <-
  tuits %>%
  
  # From "text", make rows for each string, assign the resulting words to the column "Palabra"
  unnest_tokens(input = "text", output = "Palabra") %>%
  # Make inner join with afinn lexicon in the column "Palabra"
  inner_join(afinn, ., by = "Palabra") %>%
  # Create column "Tipo" depending on "Puntuacion" as positive or negative
  mutate(Tipo = ifelse(Puntuacion > 0, "Positiva", "Negativa")) %>%
  # screen_name -> "Candidato"
  rename("Candidato" = screen_name)

head(tuits_afinn,3)
```

Now summarize the mean of these evaluations (in tuits_adinn) by candidate, then we join with the `tuits` tibble.

```{r}
tuits <-
  tuits_afinn %>%
  # Group by status id
  group_by(status_id) %>%
  # Compute the summarized mean for "Puntuacion"
  summarise(Puntuacion_tuit = mean(Puntuacion)) %>%
  # Leftjoin summarized `tuits_afinn` with `tuits` by "status_id" (tweet id) column
  # to retrieve mean for each tweet.
  left_join(tuits, ., by = "status_id") %>%
  # Avoid calculated NA values
  mutate(Puntuacion_tuit = ifelse(is.na(Puntuacion_tuit), 0, Puntuacion_tuit)) %>%
  # screen_name -> "Candidato"
  rename("Candidato" = screen_name)
head(tuits,3)
```

Total tweets by candidate

```{r}
tuits_afinn %>%
  count(Candidato) # count grouped by "Candidato"
```

# Count vocabulary used by each candidate

```{r}
tuits_afinn %>%
  group_by(Candidato) %>%
  # Unique words
  distinct(Palabra) %>%
  # Count these unique words by candidate
  count()
```

By using a function, we automate two barplots with the top 10 of both positive and negative words used by each candidate.

* We observe a high count for the word "no", so we filter the tibble to avoid these incidences:
  
```{r}
# Theme for the graphs
tema_graf <-
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "#EBEBEB", colour = NA),
        legend.position = "none",
        legend.box.background = element_rect(fill = "#EBEBEB", colour = NA))
```

```{r}
tuits_afinn <-
  tuits_afinn %>%
  filter(Palabra != "no")
```

```{r}
par(mfrow = c(2,1))

map(c("Positiva", "Negativa"), function(sentimiento) {
  tuits_afinn %>%
    filter(Tipo ==  sentimiento) %>% # Filter according to fed `sentimiento`
    
    # Count words with `sentimiento` grouped by Candidatoand sort it in descending order
    group_by(Candidato) %>%
    count(Palabra, sort = T) %>%
    
    # Take Top 10
    top_n(n = 10, wt = n) %>%
    #Plot the result for `sentimiento`
    ggplot() +
    aes(Palabra, n, fill = Candidato) +
    geom_col() +
    facet_wrap("Candidato", scales = "free") +
    scale_y_continuous(expand = c(0, 0)) +
    coord_flip() +
    labs(title = sentimiento) +
    tema_graf
})
```

Add a column with the mean evaluation for each tweet, group by `status_id`, `Candidato` and `Fecha` to summarize mean of the tweets evaluation.

```{r}
tuits_afinn_fecha <-
  tuits_afinn %>%
  group_by(status_id) %>%
  mutate(Suma = mean(Puntuacion)) %>%
  group_by(Candidato, Fecha) %>%
  summarise(Media = mean(Puntuacion))
head(tuits_afinn_fecha,3)
```

```{r}
par(mfrow = c(3,2))
#-----------------------------------------1
tuits_afinn_fecha %>%
  ggplot() +
  aes(Fecha, Media, color = Candidato) +
  geom_line() +
  ggtitle("Plot date vs. (daily) evaluation") +
  tema_graf +
  theme(legend.position = "top")
#-----------------------------------------2
tuits_afinn_fecha %>%
  ggplot() +
  aes(Fecha, Media, color = Candidato) +
  geom_hline(yintercept = 0, alpha = .35) +
  geom_line() +
  facet_grid(Candidato~.) + # Create the same plot individually for each candidate
  ggtitle("Plot date vs. (daily) evaluation by individual cadidates") +
  tema_graf + # Theme created before
  theme(legend.position = "none")
#-----------------------------------------3
tuits_afinn_fecha %>%
  ggplot() +
  aes(Fecha, Media, color = Candidato) +
  geom_smooth(method = "loess", fill = NA) +
  coord_cartesian(ylim = c(-3, 3)) +
  ggtitle("Plot date vs. (daily) evaluation with smooth lines") +
  tema_graf
#-----------------------------------------4
tuits_afinn %>%
  ggplot() +
  aes(Fecha, Puntuacion, color = Candidato) +
  geom_smooth(method = "loess", fill = NA) +
  coord_cartesian(ylim = c(-3, 3)) +
  ggtitle("Plot date vs. daily score") +
  tema_graf
#-----------------------------------------5
tuits_afinn %>%
  ggplot() +
  aes(Fecha, Puntuacion, color = Candidato) +
  geom_smooth(method = "loess", fill = NA) +
  geom_point(color = "#E5E5E5") + # To compare with individual tweets
  facet_wrap(~Candidato) +
  ggtitle("Plot date vs. daily score by individual candidates (smooth and tweets to compare)") +
  tema_graf
#-----------------------------------------6
tuits_afinn_fecha %>%
  ggplot() +
  aes(Fecha, Media, color = Candidato) +
  geom_point(color = "#E5E5E5") +
  geom_smooth(method = "lm", fill = NA) +
  facet_wrap(~Candidato) +
  ggtitle("Plot date vs. (daily) evaluation with linear regression to compare with individual tweets by candidate") +
  tema_graf
```

```{r}
par(mfrow = c(3,2))
#-----------------------------------------1
tuits_afinn_fecha %>%
  group_by(Candidato) %>%
  mutate(MediaR = rollmean(Media, k = 3, align = "right", na.pad = TRUE)) %>% # media móvil
  ggplot() +
  aes(Fecha, MediaR, color = Candidato) +
  geom_hline(yintercept = 0, alpha = .35) +
  geom_line() +
  facet_grid(Candidato~.) +
  ggtitle("Plot Date vs. Rollmean (variation from previous date) by candidate.") +
  tema_graf
#-----------------------------------------2
tuits_afinn_fecha %>%
  ggplot() +
  aes(Fecha, Media, color = Candidato) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~Candidato) +
  tema_graf
#-----------------------------------------3
# This can be improved with ggplot parameters for the proportion in geom_col, not by preliminary treatment
tuits_afinn %>%
  count(Candidato, Tipo) %>%
  group_by(Candidato) %>%
  mutate(Proporcion = n / sum(n)) %>%
  ggplot() +
  aes(Candidato, Proporcion, fill = Tipo) +
  geom_col() +
  scale_y_continuous(labels = percent_format()) +
  ggtitle("Proportional polarity by candidate") +
  tema_graf +
  theme(legend.position = "top")
#-----------------------------------------4
tuits_afinn %>%
  group_by(Candidato, Fecha) %>%
  count(Tipo) %>%
  mutate(Proporcion = n / sum(n)) %>%
  ggplot() +
  aes(Fecha, Proporcion, fill = Tipo) +
  geom_col(width = 1) +
  facet_grid(Candidato~.) +
  scale_y_continuous(labels = percent_format()) +
  scale_x_date(expand = c(0, 0)) +
  ggtitle("Proportional polarity by date for each candidate") +
  tema_graf +
  theme(legend.position = "top")
#-----------------------------------------5
tuits %>%
  ggplot() +
  aes(Candidato, Puntuacion_tuit, fill = Candidato) +
  geom_boxplot() +
  ggtitle("Boxplot of tweet scores for each candidate") +
  tema_graf
#-----------------------------------------6
tuits %>%
  mutate(Mes = factor(Mes)) %>%
  ggplot() +
  aes(Mes, Puntuacion_tuit, fill = Candidato) +
  geom_boxplot(width = 1) +
  facet_wrap(~Candidato) +
  ggtitle("Monthly score boxplots by candidate") +
  tema_graf +
  theme(legend.position = "none")
```

```{r}
par(mfrow = c(1,2))
#-----------------------------------------1
tuits %>%
  ggplot() +
  aes(Puntuacion_tuit, color = Candidato) +
  geom_density() +
  facet_wrap(~Candidato) +
  ggtitle("Scores distribution by candidate") +
  tema_graf
#-----------------------------------------2
tuits %>%
  ggplot() +
  aes(Puntuacion_tuit, color = Candidato) +
  geom_density() +
  facet_grid(Candidato~Mes) +
  ggtitle("Monthly scores distributions by Candidate") +
  tema_graf
```